---
title: "Species"
author: "Lone"
date: "2023-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(tidyverse)
library(ggplot2)
library(vegan)
library(ggmap)
library(sp)
library(reshape2)
library(VennDiagram)
library(ggvenn)
library(xlsx)
library(geosphere)

## local functions
#Create a function to generate a continuous color palette
rbPal <- colorRampPalette(c('red','blue'))

count_species <- function(row) {
  sum(row > 0)
}


```


## Load data and parse

```{r}

figs_dir = file.path("/Users/alexaei/Documents/projects/master_thesis_projects/Lone/manuscript/Fish_diversity_manuscript_figures")
data_dir = file.path("/Users/alexaei/repos/edna_citizen/data")

species_cleaned = read.csv(file.path(data_dir, "species_cleaned.csv"))
keep_samples = rownames(species_cleaned)

# get presence absence matrix
species_presence <- as.matrix((species_cleaned>0)+0)

# get number of species per sample
species_per_sample = rowSums(species_presence)
summary(species_per_sample)
```

## Merging environmental and species data
```{r}

#get data on environmental parameters
env_data <- as.data.frame(read.csv(file.path(data_dir, "environmental_data.csv"), sep = ";"))
env_data <- filter(env_data,`ID`!="NA")
rownames(env_data) = env_data$ID
env_data <- env_data[1:7]

data <- as.data.frame(read.csv(file.path(data_dir, "Data_master_project.csv")))
rownames(data) = data$ID

tmp = merge(x=env_data,y=data, by="ID", all.x=TRUE) %>%
  mutate(TN = as.numeric(TN))

# remove data corresponding to missing localities
env_data_cleaned = tmp[rownames(tmp) %in% keep_samples,]
rownames(env_data_cleaned) = env_data_cleaned$ID
env_data_cleaned = env_data_cleaned[order(rownames(env_data_cleaned)),]
groups <- env_data_cleaned$group



```

## Regression analysis latitude versus species detection

Plot regression between species observations and latitude (from outer to inner Oslofjord)

```{r}

species_cleaned$NumSpecies = species_per_sample

species_cleaned$Lat = env_data_cleaned$Lat

# Create a new variable that indicates whether the species is below or above a certain latitude (Drobaksill)
lat_group = ifelse(env_data_cleaned$Lat > 59.6678, "above", "below")
species_cleaned$lat_group = lat_group
env_data_cleaned$lat_group = lat_group

cairo_ps(file.path(figs_dir, "Fig5b_regression.eps"), width=80/25.4, 
         height=80/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
  ggplot(species_cleaned, aes(x = Lat, y = NumSpecies)) +
    geom_point() +  # Scatter plot
    geom_smooth(method = "lm", se = TRUE, color = "brown") +  # Regression line with confidence intervals
    labs(x = "Latitude",
         y = "observed species") +
    theme_classic() +
    theme(axis.text.y = element_text(size = 9),
          axis.text.x = element_text(size = 9),
          axis.title.x = element_text(size = 11), # set size of x axis label
          axis.title.y = element_text(size = 11),
          legend.text = element_text(size = 9), # set size of legend text
          legend.title = element_text(size = 11))
dev.off()

```

## Make prevalence histogram

Here we plot the histogram of number of observed species per site.

```{r}

cairo_ps(file.path(figs_dir, "S5_histogram.eps"), width=80/25.4, 
         height=80/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
  ggplot(species_cleaned, aes(x = NumSpecies)) +
    geom_histogram(binwidth = 1, fill = "#4477AA", color = "black", alpha = 0.7) +
    labs(x = "Number of species", y = "Frequency of sites") +
    theme_classic() +
    theme(axis.text.y = element_text(size = 9),
          axis.text.x = element_text(size = 9),
          axis.title.x = element_text(size = 11), # set size of x axis label
          axis.title.y = element_text(size = 11),
          legend.text = element_text(size = 9), # set size of legend text
          legend.title = element_text(size = 11))
dev.off()

```

## Beta diversity
Here we perform the beta diversity analyses
```{r}
h_nestedness = nestednodf(species_presence, order = F)

# Check distribution of residuals
pdf(file.path(figs_dir, "qqnorm.pdf"))
  qqnorm(species_presence)
dev.off()

########### Create NMDS object #############
# Standardize data

fish_mds = metaMDS(species_presence, distance = "jaccard", autotransform = FALSE, try = 200)

## subset NAs

scores_mds = scores(fish_mds)
fish_mds$stress # 0.092

res = as.data.frame(merge(scores_mds$sites,
                          env_data_cleaned,
                          by.x = "row.names",
                          by.y = "row.names"))

cairo_ps(file.path(figs_dir, "nmds_groups.eps"),
                              width=180/25.4,
                              height=180/25.4,
                              pointsize = 4,
                              bg = FALSE,
                              fallback_resolution = 300)
ggplot(res, aes(NMDS1, NMDS2)) +
       geom_point(aes(colour=group, shape = group)) +
       theme_bw(14) +
       xlab("Dimension 1") +
       ylab("Dimension 2")
dev.off()

latitude = env_data_cleaned[, "Lat"]

cairo_ps(file.path(figs_dir, "S7_nmds.eps"),
                              width=180/25.4,
                              height=180/25.4,
                              pointsize = 4,
                              bg = FALSE,
                              fallback_resolution = 300)
ggplot(res, aes(NMDS1, NMDS2)) +
       geom_point(aes(colour=Lat, shape = lat_group)) +
       theme_bw(14) +
       xlab("Dimension 1") +
       ylab("Dimension 2") +
       scale_colour_gradient(low = "#D1BBD7", high="#771111")
dev.off()


########## Permanova - adonis ########

adonis2(data.frame(species_presence) ~ groups, data = data.frame(env_data_cleaned))
adonis2(data.frame(species_presence) ~ lat_group, data = data.frame(env_data_cleaned))
adonis2(data.frame(species_presence) ~ TN, data = data.frame(env_data_cleaned))
subset_env_data = data.matrix(env_data_cleaned [,c(2:6,9)])
adonis2(data.frame(species_presence) ~ ., data = data.frame(subset_env_data))


######### Mantel tests #############

# geographic distance
geo_data = data.frame(lat  = env_data_cleaned[, "Lat"], lon  = env_data_cleaned[, "Lon"])
rownames(geo_data) = rownames(env_data_cleaned)
geo_dist = distm(geo_data)

# community distance
species_dist <- vegdist(species_presence, method = "jaccard")

# environmental distance
subset_env_data = data.matrix(env_data_cleaned [,2:6])
env_table = scale(subset_env_data, scale=T, center=T)
env_dist = vegdist(env_table, method="euclidean")

mantel.partial(species_dist, env_dist, geo_dist, method="pearson", permutations=999)
mantel.partial(species_dist, geo_dist, env_dist, method="pearson", permutations=999)



```
## Venn diagrams
Here we look into the overlaps of species in the outer and inner fjord

```{r}

```

