---
title: "Species"
author: "Lone"
date: "2023-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(tidyverse)
library(ggplot2)
library(vegan)
library(ggmap)
library(sp)
library(reshape2)
library(VennDiagram)
library("ggvenn")



```


## Load data and parse

```{r}

figs_dir = file.path("/Users/alexaei/Documents/projects/master_thesis_projects/Lone/manuscript/Fish_diversity_manuscript_figures")
data_dir = file.path("/Users/alexaei/repos/edna_citizen/data")

species <- read.csv(file.path(data_dir, "species_table_lca_n.csv"), row.names = 1)

# remove genus entries

remove_columns_without_underscore <- function(data) {
  cols_to_keep <- grepl("_", names(data))
  return(data[, cols_to_keep, drop = FALSE])
}

species <- remove_columns_without_underscore(species)

#change the name
rownames(species) <- gsub("_L001_R1_001.fastq.gz", "", rownames(species))
rownames(species) <- gsub("\\_.*", "", rownames(species))

remove_samples = c("1F", "1L", "2F", "2L", "3F", "3L", "4F", "4L")
species = species[!rownames(species) %in% remove_samples,]


#Remove all samples with less than 400 total reads
rowSums(species)
species_cleaned = species[which(rowSums(species)>400),]
# remove the positive control
species_cleaned = species_cleaned[-which(rownames(species_cleaned) == "P"),]
rowSums(species_cleaned)

#get mata data
data <- as.data.frame(read.csv(file.path(data_dir, "Data_master_project.csv")))
rownames(data) = data$ID

#remove data from samples removed in the cleaning process
keep_samples = rownames(species_cleaned)
data_cleaned = data[rownames(data) %in% keep_samples,]
data_cleaned = data_cleaned[order(rownames(data_cleaned)),]
species_cleaned = species_cleaned[order(rownames(species_cleaned)),]

#make species into proportion
prop_species_clean = t(prop.table(as.matrix(species_cleaned), margin = 1)*100)
prop_species_clean = prop_species_clean[order(rowSums(-prop_species_clean)),]
#only the 15 most species
prop_species_clean = prop_species_clean[1:15,]
melted_prop_species_clean = melt(prop_species_clean)

colnames(melted_prop_species_clean) <- c("X1", "X2", "value")

melted_prop_species_clean = dplyr::arrange(melted_prop_species_clean, X1, desc(value))
melted_prop_species_clean$X2 = factor(melted_prop_species_clean$X2, levels = unique(melted_prop_species_clean$X2))

#make color palett
color_palette_fish = c("#114477", "#4477AA", "#77AADD", "#117755",
                          "#44AA88", "#99CCBB", "#777711", "#AAAA44",
                          "#DDDD77", "#771111", "#AA4444", "#DD7777",
                          "#771144", "#AA4477", "#DD77AA")

fish_colors = setNames(color_palette_fish, levels(melted_prop_species_clean$X1))



cairo_ps(file.path(figs_dir, "S6_fish_species.eps"), width=160/25.4, height=160/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
ggplot(melted_prop_species_clean, aes(x = X2, y = value, fill = X1), cex=4)+
  geom_bar(stat = "identity", position = "stack", alpha = .5, width = 1, size = 10)+
  guides(fill = guide_legend(title = "Fish species"))+
  coord_flip() +
  xlab("Sampling points")+
  ylab("Proportion of reads [%]")+
  scale_fill_manual(values = fish_colors)+
  scale_x_discrete(limits = rev(levels(melted_prop_species_clean$X2))) +
  theme_classic() + 
  theme(axis.text = element_text(size = 9),
        axis.title.x = element_text(size = 11), # set size of x axis label
        axis.title.y = element_text(size = 11),
        legend.text = element_text(size = 11), # set size of legend text
        legend.title = element_text(size = 11))

dev.off()


```


```{r}


#get data on environmental parameters
env.data <- as.data.frame(read.csv(file.path(data_dir, "environmental_data.csv"), sep = ";"))
env.data <- filter(env.data,`ID`!="NA")
rownames(env.data) = env.data$ID

env.data <- env.data[1:7]

# remove data corresponding to missing localities
env.data_cleaned = env.data[rownames(env.data) %in% keep_samples,]
env.data_cleaned = env.data_cleaned[order(rownames(env.data_cleaned)),]
groups <- env.data_cleaned$group
env.data_cleaned <- data.frame(lapply(env.data_cleaned[1:6], as.numeric))
env.data_cleaned <- cbind(env.data_cleaned, groups)


# cairo_ps(file.path(figs_dir, "ori_map.eps"), width=80/25.4, height=160/25.4, 
#          pointsize = 4, bg = FALSE, fallback_resolution = 300)
#  ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#    geom_text(data = data, aes(x = Lon, y = Lat, label = ID), color = "red", size = 5) +
# # geom_point(data=data, aes(x=Lon, y=Lat), color="red")+
#   labs(x = "Longitude", y = "Latitude")
# dev.off()

df2 <- merge(x=env.data,y=data, by="ID", all.x=TRUE) %>%
  mutate(TN = as.numeric(TN))



#Create a function to generate a continuous color palette
rbPal <- colorRampPalette(c('red','blue'))

#This adds a column of color values
# based on the y values
df2$Col <- rbPal(10)[as.numeric(cut(df2$TN,breaks = 10))]

# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df2, aes(x=Lon, y=Lat), colour=df2$Col)+
#   labs(x = "Longitude", y = "Latitude")



```
Figure x. A map of the Oslo fjord showing the sampling points along the fjord and the nitrogen levels observed there grouped high to low, where high is … and low is represented by the color …. 


```{r}
historic.data <-  read.csv(file.path(data_dir, "historic.data.csv"), sep = ";")

historic.data <- subset(historic.data, select = c(institution, validScientificName, latitude, longitude, dateTimeCollected ))

historic.data$validScientificName <- gsub(" ", "_", historic.data$validScientificName)

colnames(historic.data) <- c("institution", "species", "Lat", "Lon", "date")

historic.data <- historic.data[grepl("_", historic.data$species), ]

# cairo_ps(file.path(figs_dir, "hist_map.eps"), width=80/25.4, 
#          height=160/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=historic.data, aes(x=longitude, y=latitude), color="red")+
#   labs(x = "Longitude", y = "Latitude")
# dev.off()
```

## Species barplot


```{r}

hal.data <- read.csv(file.path(data_dir, "hal.data.csv"), check.names = F, sep =",")

Gadus_morhua <- hal.data[, 4] + hal.data[, 7]
hal.data <- hal.data[, -7]
hal.data[, 4] <- Gadus_morhua

hall.data <- hal.data[, 3:31]
col.name <- colnames(hal.data[, 4:31])
hall.data<-hall.data %>%
  group_by(out) %>%
  dplyr::summarise(across(any_of(col.name), sum))

rownames(hall.data) <- hall.data$out

try = t(as.matrix(hall.data))

hal_melt = melt(try)

hal <- filter(hal_melt, value >0)
colnames(hal)<- c("X1","X2","value")
hal <- filter(hal, X1 != "out")


hal <- as.data.frame(subset(hal, select = c(X1, X2)))

colnames(hal) <- c("species", "out")

```

```{r}
#select only observations made after 2000
time <- as.Date(strptime(historic.data$date, "%d.%m.%Y %H:%M:%S"))
sum(time > as.Date("2017-1-1"))
historical.2000 <- historic.data[time > as.Date("2017-1-1"), ]


#make 
base.df <- subset(data_cleaned, select = c(Lon,Lat))
target.df <- subset(historical.2000, select = c(Lon, Lat))
colnames(target.df) <- c("Lon", "Lat")
target_base <- rbind(base.df, target.df)

base.mtx <- as.matrix(base.df)
target.mtx <- as.matrix(target.df)



cx <- lapply(1:nrow(target.df), function(x){
    spDistsN1(pts = base.mtx, pt = target.mtx[x, ], longlat =T)}) %>%
lapply(which.min) %>%
unlist %>%
bind_cols(target.df, out = .)

test <- plyr::match_df(cx, base.df, on=c("Lon", "Lat"))

cy <- lapply(1:nrow(target.df), function(x){
    spDistsN1(pts = base.mtx, pt = target.mtx[x, ], longlat =T)}) %>%
lapply(min) %>%
unlist %>%
bind_cols(target.df, distance = .)

cw <- cbind(cx, cy)
cw <- cw[, !duplicated(colnames(cw))]
cw <- cbind(cw, historical.2000$species)
cw <- filter(cw, distance<1)


spc <- species_cleaned
rownames(spc)<- 1:79

ps = t(prop.table(as.matrix(spc), margin = 1)*100)

mps = melt(ps)

original <- filter(mps, value >0)

colnames(original)<- c("X1","X2","value")
original <- as.data.frame(subset(original, select = c(X1, X2)))

colnames(original) <- c("species", "out")

colnames(cw) <- c("Lon", "Lat", "out", "distance", "species")
cc <- subset(cw, select = c(out, species))
cc <- rbind(cc, hal)

cc$species <- gsub(" ", "_", cc$species)
colnames(cc) <- c( "out", "species")

#cc$species <- gsub("\\_.*", "", rownames(species_cleaned))

test2 <- plyr::match_df(cc, original, on = c("species", "out"))
test2 <- unique(test2$species)

values <- c(1:79)
names <- rownames(base.df)

#for (y in 1:length(values)){
 # for (i in 1:length(test2$out)){
  #if (test2$out[i] == values[y]){
   # dx <- replace(test2$out, i, names[y])
  #}
#}
#}

all_species <- as.data.frame(unique(historic.data$species))
colnames(all_species) <- "species"
hal_species <- as.data.frame(unique(hal$species))
colnames(hal_species) <- "species"
colnames(all_species) <- "species"
all_species <- rbind(all_species, hal_species)
all_species <- as.data.frame(unique(all_species))
recent_species <- as.data.frame(unique(historical.2000$species))
colnames(recent_species) <- "species"
colnames(recent_species) <- "species"

recent_species <- rbind(recent_species, hal_species)
recent_species <- as.data.frame(unique(recent_species))
lost_species <- rbind(recent_species,all_species)

lost_species <- as.data.frame(table(lost_species))

lost_species <- subset(lost_species,Freq==1)
lost_species <- as.data.frame(lost_species[,-2])
colnames(lost_species) <- "species"

my_species <- as.data.frame(unique(original$species))
colnames(my_species) <- "species"

new_species <- rbind(my_species,lost_species)

new_species <- as.data.frame(table(new_species))

refound <- subset(new_species,Freq==2)
refound <- as.data.frame(refound[,-2])
colnames(refound) <- "species"


lost <- rbind(lost_species, refound)
lost <- as.data.frame(table(lost))
lost <- subset(lost,Freq==1)

#if i need to make observation data in columns
#library(tidyr)
#historical.2000_new <- historical.2000 %>% pivot_wider(names_from = validScientificName, values_from = longitude)

recent_species$type <- "Recent species"
colnames(recent_species) <- c("species", "type")

hist_species <- as.data.frame(unique(cc$species))
hist_species$type <- "Historic species"
colnames(hist_species) <- c("species", "type")


out_of_reach <- rbind(recent_species, hist_species)
out_of_reach <- as.data.frame(table(out_of_reach))
out_of_reach <- subset(out_of_reach,Freq==0)
out_of_reach$type <- "Out of reach"
colnames(out_of_reach) <- c("species", "type")

lost <- subset(lost, select = species)
lost$type <- "lost"                           

blebb <- plyr::match_df(recent_species, cc, on = "species")
blebb <- unique(blebb$species)
bebe <- unique(cc$species)
babe <- unique(recent_species$species)

```
## comparisons of method outputs

add sample IDs to identify which points are closest
and run fisher.test and chisq.test
```{r}

# Grouping your dataframe based on species and then taking the number of observations in each group

df_historic_summary <-  cc %>%
  group_by(species) %>%
  dplyr::summarize(n= length(species))

df_historic_summary$type <- "Traditional method"

# Doing the same with your data frame. There's something fishy going on in the df_lone: Namely it seems that all species are present in all sites, and this can't be true, so I recommend you to check this dataframe?
df_lone_summary <-  original %>%
  group_by(species) %>%
  dplyr::summarize(n= length(species))

df_lone_summary$type <- "eDNA method"

#Combining the dataframes

df_both_summary <- rbind(df_historic_summary,df_lone_summary)


cairo_ps(file.path(figs_dir, "Fig2_fish_species.eps"), width=80/25.4, 
         height=160/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
# Here's the very basic plot but you can make it prettier i.e. select only species of interest, change color scheme, perhaps add the percentage number value on the corresponding box etc.
ggplot(df_both_summary, aes(type, species, fill= n)) + 
  geom_tile(alpha = 0.7) +
  labs(fill = "") +
  theme_classic() + 
    theme(axis.text.y = element_text(size = 6),
          axis.text.x = element_text(size = 8, angle = 15, hjust = 1),
          axis.title.x = element_text(size = 11), # set size of x axis label
          axis.title.y = element_text(size = 11),
          legend.text = element_text(size = 9), # set size of legend text
          legend.title = element_text(size = 11))
# I thin. k it would be neat if you can turn the values into relative abundance. So basically instead of showing that cod was present in 5 sampling sites, you show that e.g. cod was present in 8% of the sampling sites.

dev.off()

lone_summary <- subset(df_lone_summary, select = c(species, type))
historic_summary <- subset(df_historic_summary, select = c(species, type))


match_my <- plyr::match_df(historic_summary, lone_summary, on= "species")
match_his <- plyr::match_df(out_of_reach, lone_summary, on= "species")
cc_species <- as.data.frame(unique(cc$species))
colnames(cc_species) <- "species"
hal_species2 <- as.data.frame(unique(hal$species))
colnames(hal_species2) <- "species"
mach <- plyr::match_df(cc_species,hal_species2 , on ="species")

no_species <- rbind(match_my,historic_summary)

no_species <- as.data.frame(table(no_species))

no_species <- subset(no_species,Freq==1)
no_species <- as.data.frame(no_species[,-2])
colnames(no_species) <- "species"

makka <- plyr::match_df(no_species, historic_summary)

```

## Detailed analyses 

Here we check which species are outside the coordinates. Particularly we are interested in species that are found in the center of the fjord - non littoral.  

We also look into individual species and rare species, as well as warm water species.


```{r}


df_historic <- cc
df_lone <- original

df_lone$type <- "eDNA method"
df_historic$type <- "Traditional method"

df_historic <- unique(df_historic)

#hist(as.numeric(df_historic$out), breaks = 80)

df_both_map <- rbind(df_lone,df_historic)

#	Let's use Gadus_morhua as an example species. You can repeat this same code for all the species that you are interested in by just assigning a different species below:

genus_species <- "Gadus_morhua"

# Here you subset the data for only the species of interest
df_genus_species <- subset(df_both_map,species==genus_species)

#Here you create a dataframe that counts the number of observations at each sampling point
df_genus_species2 <- df_genus_species %>%
  group_by(out) %>%
  dplyr::summarize(n= length(out))

# Here you merge the "species of intersest" data frame and the number of observations at each sampling point
df_genus_species <- merge(df_genus_species,df_genus_species2,by="out")

# Here you create a new categorical variable indicating whether the species of interest was present in both the historical record and your sampling, or only historical record or your sampling.
df_genus_species$presence <- ifelse(df_genus_species$n>1,"Present in both",df_genus_species$type)

# Here you create an index dataframe for all those sampling sites where this species was present
sites_species <- as.data.frame(unique(df_genus_species$out))
colnames(sites_species)[1] ="out"

#Here you create an index dataframe for all those sampling sites where this species was absent
all_sites <- as.data.frame(1:79)
colnames(all_sites)[1] ="out"
sites_empty <- rbind(sites_species,all_sites)



#sites_empty <- sites_empty %>%
  #group_by(out) %>%
  #mutate(out2 = ifelse(length(out)>1,1,0))

sites_empty <- as.data.frame(table(sites_empty))



sites_empty <- subset(sites_empty,Freq==1)
sites_empty$species <- genus_species
sites_empty$type <- NA
sites_empty$n <- 0
sites_empty$presence <-"Absent" # Here you finally assign "absent" value to those sites where this species was absent in both the historical record and your data.
sites_empty <- sites_empty[,-2]

#Here you make the final dataframe that you can use in your maps.

df_genus_species <- rbind(df_genus_species,sites_empty)

df_genus_species <- df_genus_species[,c(1,2,5)]

df_genus_species <- df_genus_species[!duplicated(df_genus_species), ]

#rownames(df_genus_species) <- df_genus_species$out)

coordinates <- subset(data_cleaned, select = c(Lat,Lon))
coordinates$out <- 1:79


df_genus_species <- merge(df_genus_species, coordinates, by = "out")

set_shapes <- setNames(c(18, 15, 17, 16), levels(unique(df_genus_species$presence)))
# Now you can just add the coordinates for each sampling site and make your map figure. In order to get the different symbols to represent the different presence status, just add "presence" column as a grouping variable.
# png("cod-map.png", height = 500, width = 400)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df_genus_species, aes(x=Lon, y=Lat, color = presence, shape=presence)) +
#   scale_shape_manual(values = set_shapes) +
#   labs(x = "Longitude", y = "Latitude")+
#   theme(axis.text = element_text(size = 14),
#         axis.title.x = element_text(size = 16), # set size of x axis label
#         axis.title.y = element_text(size = 16),
#         legend.text = element_text(size = 14), # set size of legend text
#         legend.title = element_text(size = 16))
# dev.off()
```
```{r}
genus_species <- "Ctenolabrus_rupestris"

# Here you subset the data for only the species of interest
df_genus_species <- subset(df_both_map,species==genus_species)

#Here you create a dataframe that counts the number of observations at each sampling point
df_genus_species2 <- df_genus_species %>%
  group_by(out) %>%
  dplyr::summarize(n= length(out))

# Here you merge the "species of intersest" data frame and the number of observations at each sampling point
df_genus_species <- merge(df_genus_species,df_genus_species2,by="out")

# Here you create a new categorical variable indicating whether the species of interest was present in both the historical record and your sampling, or only historical record or your sampling.
df_genus_species$presence <- ifelse(df_genus_species$n>1,"Present in both",df_genus_species$type)

# Here you create an index dataframe for all those sampling sites where this species was present
sites_species <- as.data.frame(unique(df_genus_species$out))
colnames(sites_species)[1] ="out"

#Here you create an index dataframe for all those sampling sites where this species was absent
all_sites <- as.data.frame(1:79)
colnames(all_sites)[1] ="out"
sites_empty <- rbind(sites_species,all_sites)


#sites_empty <- sites_empty %>%
  #group_by(out) %>%
  #mutate(out2 = ifelse(length(out)>1,1,0))

sites_empty <- as.data.frame(table(sites_empty))
sites_empty <- subset(sites_empty,Freq==1)
sites_empty$species <- genus_species
sites_empty$type <- NA
sites_empty$n <- 0
sites_empty$presence <-"Absent" # Here you finally assign "absent" value to those sites where this species was absent in both the historical record and your data.
sites_empty <- sites_empty[,-2]

#Here you make the final dataframe that you can use in your maps.
df_genus_species <- rbind(df_genus_species,sites_empty)
df_genus_species <- df_genus_species[,c(1,2,5)]
df_genus_species <- df_genus_species[!duplicated(df_genus_species), ]

#rownames(df_genus_species) <- as.numeric(df_genus_species$out)
coordinates <- subset(data_cleaned, select = c(Lat,Lon))
coordinates$out <- 1:79

df_genus_species <- merge(df_genus_species, coordinates, by = "out")

# Now you can just add the coordinates for each sampling site and make your map figure. In order to get the different symbols to represent the different presence status, just add "presence" column as a grouping variable.
# png("bergnebb-map.png", height = 500, width = 400)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df_genus_species, aes(x=Lon, y=Lat, color = presence, shape=presence)) +
#   labs(x = "Longitude", y = "Latitude")+
#   scale_shape_manual(values = set_shapes) +
#   theme(axis.text = element_text(size = 14),
#         axis.title.x = element_text(size = 16), # set size of x axis label
#         axis.title.y = element_text(size = 16),
#         legend.text = element_text(size = 14), # set size of legend text
#         legend.title = element_text(size = 16))
# dev.off()
```
```{r}


df2 <- subset(df_both_map, species %in% c("Oncorhynchus_gorbuscha", "Cyprinus_carpio", "Tinca_tinca", "Carassius_auratus"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
# png("invasive-map.png", height = 400, width = 400)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
#   labs(x = "Longitude", y = "Latitude")+
#   scale_shape_manual(values = c(15, 16)) +
#   theme(axis.text = element_text(size = 14),
#         axis.title.x = element_text(size = 16), # set size of x axis label
#         axis.title.y = element_text(size = 16),
#         legend.text = element_text(size = 14), # set size of legend text
#         legend.title = element_text(size = 16))
# dev.off()

df2 <- subset(df_both_map, species %in% c("Engraulis_encrasicolus", "Scomber_scombrus", "Entelurus_aequoreus", "Trisopterus_minutus", "Sardina_pilchardus"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
png("warm-water-map.png", height = 400, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = c(15, 16)) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()
```
```{r}
# Basic piechart, må fikse
df <- as.data.frame(table(historical.2000$institution))

colnames(df) <- c("Institution", "frequency")

#eller simpelthen bare
png("org_barplot.png", height = 300, width = 450)
ggplot(df, aes(x="", y=frequency, fill=Institution)) +
  theme_classic() + 
 geom_bar(stat="identity")
dev.off()
```
```{r}
df2 <- subset(df_both_map, species %in% c("Buglossidium_luteum", "Ciliata_septentrionalis"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
# png("novel-map.png", height = 400, width = 400)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=species)) +
#   labs(x = "Longitude", y = "Latitude")+
#   scale_shape_manual(values = c(15, 16)) +
#   theme(axis.text = element_text(size = 14),
#         axis.title.x = element_text(size = 16), # set size of x axis label
#         axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), 
## set size of legend text
#         legend.title = element_text(size = 16))
# dev.off()
```

```{r}

df2rare <- subset(df_both_map, species %in% c("Trisopterus_luscus", "Phrynorhombus_norvegicus", "Ammodytes_marinus", "Raniceps_raninus", "Centrolabrus_exoletus", "Enchelyopus_cimbrius"))
df2rare$species <- factor(df2rare$species)

df3 <- merge(df2rare, coordinates)
# png("rare-map.png", height = 500, width = 450)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
#   labs(x = "Longitude", y = "Latitude")+
#   scale_shape_manual(values = c(15, 16)) +
#   theme(axis.text = element_text(size = 14),
#         axis.title.x = element_text(size = 16), # set size of x axis label
#         axis.title.y = element_text(size = 16),
#         legend.text = element_text(size = 14), # set size of legend text
#         legend.title = element_text(size = 16))
# dev.off()

df2 <- subset(df_both_map, species %in% refound$species)
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
# png("refound-map.png", height = 400, width = 400)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df3, aes(x=Lon, y=Lat, color = species)) +
#   labs(x = "Longitude", y = "Latitude")+
#   scale_shape_manual(values = c(15, 16)) +
#   theme(axis.text = element_text(size = 14),
#         axis.title.x = element_text(size = 16), # set size of x axis label
#         axis.title.y = element_text(size = 16),
#         legend.text = element_text(size = 14), # set size of legend text
#         legend.title = element_text(size = 16))
# dev.off()

df2 <- subset(df_both_map, species %in% c("Gymnocephalus_cernua", "Entelurus_aequoreus", "Abramis_brama", "Phrynorhombus_norvegicus"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
# ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
#   geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
#   labs(x = "Longitude", y = "Latitude")

```

## Comparisons of methods

Fisher exact and chisq tests




## The statistical analyses

```{r}

#Ditta går ikkje når 
c_species <- gsub(" ", "_", cc$species)
c_species <- unique(c_species)
o_species <- as.character(unique(original$species))
species_name <- c(c_species, o_species)
species_name <- unique(species_name)


his2017 <- subset(historical.2000, select = c("species", "Lon", "Lat"))

mx3 = cc
mx4 = original
sn <- unique(cc$species)
ft.p <- c()
ct.p <- c()

for (specie in sn){
  subset_cc = mx3[mx3$species == specie,]
  subset_og = mx4[mx4$species == specie,]
  merged_species = merge(subset_cc, subset_og, by = "out", all = TRUE)
  merged_species$species.y = as.character(merged_species$species.y)
  merged_species[is.na(merged_species)] <- 0
  if(length(unique(merged_species$species.x))>1 &&
     length(unique(merged_species$species.y))>1){

    ct <- chisq.test(merged_species$species.x, merged_species$species.y, simulate.p.value = T)
    ft <- fisher.test(merged_species$species.x, merged_species$species.y, simulate.p.value = T)
    ft.p <- append(ft.p, ft$p.value)
    ct.p <- append(ct.p, ct$p.value)
    }
  else{
    ft.p <- append(ft.p, NA)
    ct.p <- append(ct.p, NA)
  }  
}
names(ft.p) = sn
names(ct.p) = sn



ft.p.adjust <- p.adjust(ft.p, method = "hochberg", n = length(ft.p))
ct.p.adjust <- p.adjust(ct.p, method = "hochberg", n = length(ct.p))

ft.p.adjust <- as.data.frame(ft.p.adjust)
ct.p.adjust <- as.data.frame(ct.p.adjust)

both.test <- as.data.frame(cbind(ft.p.adjust, ct.p.adjust))
both.test$species  <- rownames(both.test)
colnames(both.test) <- c("Fisher test","Chi-squared test", "species")

both.test <- na.omit(both.test)

both.test <- melt(both.test)

# Create a binary variable indicating whether each value is below 0.1
both.test$below_005 <- ifelse(both.test$value < 0.05, "below 0.05", ifelse(both.test$value < 0.1, "below 0.1", "above 0.1"))

# Plot the heatmap with a manual fill scale

my_breaks = c(0.01, 0.05, 0.001, 0.0001)
ggplot(both.test, aes(variable, species, fill = below_005)) +
  geom_tile() +
  scale_fill_manual(values = c("dark blue", "blue", "light blue"), 
                    labels = c("above 0.1", "below 0.1", "below 0.05"))


cairo_ps(file.path(figs_dir, "Fig3_stats_test.eps"), width=90/25.4, 
         height=160/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
ggplot(both.test, aes(variable, species, fill = below_005)) +  
  geom_tile(alpha = 0.5) + 
  scale_fill_manual(values = c("dark blue", "blue", "light blue"), 
                    labels = c("> 0.1", "< 0.1", "< 0.05")) +
  theme_classic() + 
    theme(axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8, angle = 15, hjust = 1),
          axis.title.x = element_text(size = 11), # set size of x axis label
          axis.title.y = element_text(size = 11),
          legend.text = element_text(size = 9), # set size of legend text
          legend.title = element_text(size = 11))
dev.off()



```

## Venn diagram

Plotting species presence and overlap between inner and outer Oslofjord

```{r}

# Create a new variable that indicates whether the species is below or above a certain latitude
lat_group <- ifelse(data_cleaned$Lat > 59.667800, "above", "below")

df4 <- data.frame(lat_group, species_cleaned)

melted_df= melt(df4)
melt_df <- filter(melted_df, value !=0)
melt_df <- subset(melt_df, select = -c(value))

above = c()
below = c()
for (i in 1:length(melt_df$lat_group)){
  if (melt_df$lat_group[i] =="above"){
  above <- melt_df$variable[i]
  }
  else{below <- melt_df$variable[i]}
}

# use list as input
A <-list("Outer Oslo fjord" = melt_df[melt_df$lat_group == "below", "variable"],
           "Inner Oslo fjord" = melt_df[melt_df$lat_group == "above", "variable"])

above <- as.data.frame(A$`Inner Oslo fjord`)
above <- unique(above)
colnames(above) <- "species"
below <- as.data.frame(A$`Outer Oslo fjord`)
below <- unique(below)
colnames(below) <- "species"

bee <- rbind(above, below)
bee <- as.data.frame(table(bee))
bee <- subset(bee,Freq==1)
bee <- as.data.frame(bee)
colnames(bee) <- c("species", "v1")

abee <- plyr::match_df(above, bee, on = "species")
bbee <- plyr::match_df(below, bee, on = "species")

spge <- rbind(abee, bbee)

# create venn diagram and display all the sets
cairo_ps(file.path(figs_dir, "Fig5a_venn.eps"), width=80/25.4, 
         height=80/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
ggvenn(A) +
     theme(
        legend.text = element_text(size = 11), # set size of legend text
        legend.title = element_text(size = 11),
        plot.subtitle = element_text(size = 11), 
        text = element_text(size = 11)) # set the font size to 18
dev.off()
```

## Regression analysis latitude versus species detection

Plot regression between species observations and latitude (from outer to inner Oslofjord)

```{r}

count_species <- function(row) {
  sum(row > 0)
}

species_cleaned$NumSpecies <- apply(species_cleaned[, -1], 1, count_species)

species_cleaned$Lat = data_cleaned$Lat

species_cleaned$lat_group = lat_group

cairo_ps(file.path(figs_dir, "Fig5b_regression.eps"), width=80/25.4, 
         height=80/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
ggplot(species_cleaned, aes(x = Lat, y = NumSpecies)) +
  geom_point() +  # Scatter plot
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # Regression line with confidence intervals
  labs(x = "Latitude",
       y = "observed species") +
  theme_classic() +
  theme(axis.text.y = element_text(size = 9),
        axis.text.x = element_text(size = 9),
        axis.title.x = element_text(size = 11), # set size of x axis label
        axis.title.y = element_text(size = 11),
        legend.text = element_text(size = 9), # set size of legend text
        legend.title = element_text(size = 11))
dev.off()

```

## Make prevalence histogram

Here we plot the histogram of number of observed species per site.

```{r}

```

