---
title: "Species"
author: "Lone"
date: "2023-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and parse

```{r}
library(tidyverse)
library(reshape)

species <- read.csv("~/Desktop/Master/Analyses/species_table.csv", row.names = 1)

#change the name
rownames(species) <- gsub("_L001_R1_001.fastq.gz", "", rownames(species))
rownames(species) <- gsub("\\_.*", "", rownames(species))


#Remove all samples with less than 400 total reads
rowSums(species)
species_cleaned = species[which(rowSums(species)>400),]
# remove the positive control
species_cleaned = species_cleaned[-80,]


#make species into proportion
prop_species_clean = t(prop.table(as.matrix(species_cleaned), margin = 1)*100)
prop_species_clean = prop_species_clean[order(rowSums(-prop_species_clean)),]
#only the 15 most species
prop_species_clean = prop_species_clean[1:15,]
melted_prop_species_clean = melt(prop_species_clean)

colnames(melted_prop_species_clean) <- c("X1", "X2", "value")

melted_prop_species_clean = dplyr::arrange(melted_prop_species_clean, X1, desc(value))
melted_prop_species_clean$X2 = factor(melted_prop_species_clean$X2, levels = unique(melted_prop_species_clean$X2))

#make color palett
color_palette_fish = c("#114477", "#4477AA", "#77AADD", "#117755",
                          "#44AA88", "#99CCBB", "#777711", "#AAAA44",
                          "#DDDD77", "#771111", "#AA4444", "#DD7777",
                          "#771144", "#AA4477", "#DD77AA")

fish_colors = setNames(color_palette_fish, levels(melted_prop_species_clean$X1))



cairo_ps(file.path(figs_dir, "fish_species.eps"), width=80/25.4, height=80/25.4, pointsize = 4, bg = FALSE, fallback_resolution = 300)
ggplot(melted_prop_species_clean, aes(x = X2, y = value, fill = X1), cex=4)+
  geom_bar(stat = "identity", position = "stack", alpha = .5, width = 1, size = 10)+
  guides(fill = guide_legend(title = "Fish species"))+
  coord_flip() +
  xlab("Sampling points")+
  ylab("Proportion of reads [%]")+
  scale_fill_manual(values = fish_colors)+
  scale_x_discrete(limits = rev(levels(melted_prop_species_clean$X2))) +
  theme_classic() + 
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))

dev.off()


```
samme plott men uten dei sampling point med under 400 arta

```{r}


```


```{r}
#get meta data
data <- as.data.frame(read.csv("~/Desktop/Master/Analyses/Data_master_project.csv"))
rownames(data) = data$ID


rows_to_keep <- c("11", "12", "18", "23", "45", "48", "79", "86", "91", "96", "95", "98", "111", "120", "139", "146", "160", "162", "201", "205")
bac <- data[rownames(data) %in% rows_to_keep, ]

#remove data from samples removed in the cleaning process
keep_samples = rownames(species_cleaned)
data_cleaned = data[rownames(data) %in% keep_samples,]
data_cleaned = data_cleaned[order(rownames(data_cleaned)),]
species_cleaned = species_cleaned[order(rownames(species_cleaned)),]

library(vegan)
library(ggplot2)


#get data on environmental perimeters
env.data <- as.data.frame(read.csv("~/Desktop/Master/Analyses/environmental_data.csv", sep = ";"))
env.data <- filter(env.data,`ID`!="NA")
rownames(env.data) = env.data$ID

env.data <- env.data[1:7]

# remove data corresponding to missing localities
env.data_cleaned = env.data[rownames(env.data) %in% keep_samples,]
env.data_cleaned = env.data_cleaned[order(rownames(env.data_cleaned)),]
groups <- env.data_cleaned$group
env.data_cleaned <- data.frame(lapply(env.data_cleaned[1:6], as.numeric))
env.data_cleaned <- cbind(env.data_cleaned, groups)


## include absorbance data

#create a list of the files from your target directory
file_list <- list.files("~/Desktop/Master/Lone")

#add all files into dataset
df <- NULL
for (i in 1:length(file_list)) {
  di <- read.table(paste("~/Desktop/Master/Lone", file_list[i], sep="/"), sep = ",", skip=2)
  
  df <- rbind(df, data.frame(id=gsub(".txt", "", file_list[i]), wl=di$V1, abs=di$V2))
}

#add id
df$id <- as.numeric(substr(df$id, 7,9))
df <- subset(df, complete.cases(df)) # Remove NAs

library(ggplot2)

df.400 <- subset(df, wl == 400)

df.700 <- subset(df, wl == 700)




#filtering out data to lost localities
keep_samples <- as.numeric(keep_samples)

rownames(df.400) = df.400$id
df.400_cleaned = df.400[rownames(df.400) %in% keep_samples,]
df.400_cleaned = df.400_cleaned[order(rownames(df.400_cleaned)),]

rownames(df.700) = df.700$id
df.700_cleaned = df.700[rownames(df.700) %in% keep_samples,]
df.700_cleaned = df.700_cleaned[order(rownames(df.700_cleaned)),]

#make data point for all missing locations for 400 and 700nm wavelengths
y <- env.data_cleaned$ID[!(env.data_cleaned$ID %in% df.400_cleaned$id)]

for (i in 1:length(y)){
  df.400_cleaned[nrow(df.400_cleaned) + 1,] <- c(y[i], 400, NA)

}

for (i in 1:length(y)){
  df.700_cleaned[nrow(df.700_cleaned) + 1,] <- c(y[i], 700, NA)

}

#install.packages("mice")
library(mice)
#library(VIM)
#plot showing how much data must be implemneted?
#aggr_plot <- aggr(df.400_cleaned, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(df.400_cleaned), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
#impliment data for 400 and 700 wavelengths
tempData <- mice(df.400_cleaned,m=5,maxit=50,meth='pmm',seed=500)
new.400 <- complete(tempData, 1)

tempData <- mice(df.700_cleaned,m=5,maxit=50,meth='pmm',seed=500)
new.700 <- complete(tempData, 1)

#remove background noice from 400 wavelength dataset
Absorbance= new.400$abs - new.700$abs
# add to environmental parameter data
env.data_cleaned <- cbind(env.data_cleaned, Absorbance)


library(ggmap)

png("ori_map.png", height = 1500, width = 1200)
 ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
   geom_text(data = data, aes(x = Lon, y = Lat, label = ID), color = "red", size = 5) +
# geom_point(data=data, aes(x=Lon, y=Lat), color="red")+
  labs(x = "Longitude", y = "Latitude")
dev.off()
df2 <- merge(x=env.data,y=data, by="ID", all.x=TRUE) %>%
  mutate(TN = as.numeric(TN))



#Create a function to generate a continuous color palette
rbPal <- colorRampPalette(c('red','blue'))

#This adds a column of color values
# based on the y values
df2$Col <- rbPal(10)[as.numeric(cut(df2$TN,breaks = 10))]

ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df2, aes(x=Lon, y=Lat), colour=df2$Col)+
  labs(x = "Longitude", y = "Latitude")



```
Figure x. A map of the Oslo fjord showing the sampling points along the fjord and the nitrogen levels observed there grouped high to low, where high is … and low is represented by the color …. 





euclidean distance

how much is due to dispersal of dna 

add numbers 1 or 0 for precense in map
```{r}
historic.data <-  read.csv("~/Desktop/Master/Analyses/historic.data.csv", sep = ";")

historic.data <- subset(historic.data, select = c(institution, validScientificName, latitude, longitude, dateTimeCollected ))

historic.data$validScientificName <- gsub(" ", "_", historic.data$validScientificName)

colnames(historic.data) <- c("institution", "species", "Lat", "Lon", "date")

historic.data <- historic.data[grepl("_", historic.data$species), ]



#install.packages("ggmap")
library(ggmap)

ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=historic.data, aes(x=longitude, y=latitude), color="red")+
  labs(x = "Longitude", y = "Latitude")

```
barplott med art mot area?
```{r}


library(sp)
#library(dplyr)
#library(purrr)
library(plyr)


hal.data <- read.csv("~/Desktop/Master/Analyses/hal.data.csv", check.names = F, sep =";")

Gadus_morhua <- hal.data[, 4] + hal.data[, 7]
hal.data <- hal.data[, -7]
hal.data[, 4] <- Gadus_morhua

hall.data <- hal.data[, 3:31]
col.name <- colnames(hal.data[, 4:31])
hall.data<-hall.data %>%
  group_by(out) %>%
  dplyr::summarise(across(any_of(col.name), sum))

rownames(hall.data) <- hall.data$out

try = t(as.matrix(hall.data))

hal_melt = melt(try)

hal <- filter(hal_melt, value >0)
colnames(hal)<- c("X1","X2","value")
hal <- filter(hal, X1 != "out")


hal <- as.data.frame(subset(hal, select = c(X1, X2)))

colnames(hal) <- c("species", "out")

```

```{r}
#select only observations made after 2000
time <- as.Date(strptime(historic.data$date, "%d.%m.%Y %H:%M:%S"))
sum(time > as.Date("2017-1-1"))
historical.2000 <- historic.data[time > as.Date("2017-1-1"), ]


#make 
base.df <- subset(data_cleaned, select = c(Lon,Lat))
target.df <- subset(historical.2000, select = c(Lon, Lat))
colnames(target.df) <- c("Lon", "Lat")
target_base <- rbind(base.df, target.df)




base.mtx <- as.matrix(base.df)
target.mtx <- as.matrix(target.df)



cx <- lapply(1:nrow(target.df), function(x){
    spDistsN1(pts = base.mtx, pt = target.mtx[x, ], longlat =T)}) %>%
lapply(which.min) %>%
unlist %>%
bind_cols(target.df, out = .)

test <- match_df(cx, base.df, on=c("Lon", "Lat"))

cy <- lapply(1:nrow(target.df), function(x){
    spDistsN1(pts = base.mtx, pt = target.mtx[x, ], longlat =T)}) %>%
lapply(min) %>%
unlist %>%
bind_cols(target.df, distance = .)

cw <- cbind(cx, cy)
cw <- cw[, !duplicated(colnames(cw))]
cw <- cbind(cw, historical.2000$species)
cw <- filter(cw, distance<1)


spc <- species_cleaned
rownames(spc)<- 1:79

ps = t(prop.table(as.matrix(spc), margin = 1)*100)

mps = melt(ps)

original <- filter(mps, value >0)

colnames(original)<- c("X1","X2","value")
original <- as.data.frame(subset(original, select = c(X1, X2)))

colnames(original) <- c("species", "out")

colnames(cw) <- c("Lon", "Lat", "out", "distance", "species")
cc <- subset(cw, select = c(out, species))
cc <- rbind(cc, hal)

cc$species <- gsub(" ", "_", cc$species)
colnames(cc) <- c( "out", "species")

#cc$species <- gsub("\\_.*", "", rownames(species_cleaned))

test2 <- match_df(cc, original, on = c("species", "out"))
test2 <- unique(test2$species)

values <- c(1:79)
names <- rownames(base.df)

#for (y in 1:length(values)){
 # for (i in 1:length(test2$out)){
  #if (test2$out[i] == values[y]){
   # dx <- replace(test2$out, i, names[y])
  #}
#}
#}

all_species <- as.data.frame(unique(historic.data$species))
colnames(all_species) <- "species"
hal_species <- as.data.frame(unique(hal$species))
colnames(hal_species) <- "species"
colnames(all_species) <- "species"
all_species <- rbind(all_species, hal_species)
all_species <- as.data.frame(unique(all_species))
recent_species <- as.data.frame(unique(historical.2000$species))
colnames(recent_species) <- "species"
colnames(recent_species) <- "species"

recent_species <- rbind(recent_species, hal_species)
recent_species <- as.data.frame(unique(recent_species))
lost_species <- rbind(recent_species,all_species)

lost_species <- as.data.frame(table(lost_species))

lost_species <- subset(lost_species,Freq==1)
lost_species <- as.data.frame(lost_species[,-2])
colnames(lost_species) <- "species"

my_species <- as.data.frame(unique(original$species))
colnames(my_species) <- "species"

new_species <- rbind(my_species,lost_species)

new_species <- as.data.frame(table(new_species))

refound <- subset(new_species,Freq==2)
refound <- as.data.frame(refound[,-2])
colnames(refound) <- "species"


lost <- rbind(lost_species, refound)
lost <- as.data.frame(table(lost))
lost <- subset(lost,Freq==1)

#if i need to make observation data in columns
#library(tidyr)
#historical.2000_new <- historical.2000 %>% pivot_wider(names_from = validScientificName, values_from = longitude)

recent_species$type <- "Recent species"
colnames(recent_species) <- c("species", "type")

hist_species <- as.data.frame(unique(cc$species))
hist_species$type <- "Historic species"
colnames(hist_species) <- c("species", "type")


out_of_reach <- rbind(recent_species, hist_species)
out_of_reach <- as.data.frame(table(out_of_reach))
out_of_reach <- subset(out_of_reach,Freq==0)
out_of_reach$type <- "Out of reach"
colnames(out_of_reach) <- c("species", "type")

lost <- subset(lost, select = species)
lost$type <- "lost"                           

blebb <- match_df(recent_species, cc, on = "species")
blebb <- unique(blebb$species)
bebe <- unique(cc$species)
babe <- unique(recent_species$species)

```
legg til sampleid så eg veit hvilket punkt det er nærmest


fisher.test og chisq.test
```{r}

# Grouping your dataframe based on species and then taking the number of observations in each group

df_historic_summary <-  cc %>%
  group_by(species) %>%
  dplyr::summarize(n= length(species))

df_historic_summary$type <- "Traditional method"

# Doing the same with your data frame. There's something fishy going on in the df_lone: Namely it seems that all species are present in all sites, and this can't be true, so I recommend you to check this dataframe?
df_lone_summary <-  original %>%
  group_by(species) %>%
  dplyr::summarize(n= length(species))

df_lone_summary$type <- "eDNA method"

#Combining the dataframes

df_both_summary <- rbind(df_historic_summary,df_lone_summary)


png("species_detected.png", height = 1000, width = 700)
# Here's the very basic plot but you can make it prettier i.e. select only species of interest, change color scheme, perhaps add the percentage number value on the corresponding box etc.
ggplot(df_both_summary, aes(type, species, fill= n)) + 
  geom_tile()+
  theme_classic() + 
    theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
# I think it would be neat if you can turn the values into relative abundance. So basically instead of showing that cod was present in 5 sampling sites, you show that e.g. cod was present in 8% of the sampling sites.

dev.off()

lone_summary <- subset(df_lone_summary, select = c(species, type))
historic_summary <- subset(df_historic_summary, select = c(species, type))


match_my <- match_df(historic_summary, lone_summary, on= "species")
match_his <- match_df(out_of_reach, lone_summary, on= "species")
cc_species <- as.data.frame(unique(cc$species))
colnames(cc_species) <- "species"
hal_species2 <- as.data.frame(unique(hal$species))
colnames(hal_species2) <- "species"
mach <- match_df(cc_species,hal_species2 , on ="species")

no_species <- rbind(match_my,historic_summary)

no_species <- as.data.frame(table(no_species))

no_species <- subset(no_species,Freq==1)
no_species <- as.data.frame(no_species[,-2])
colnames(no_species) <- "species"

makka <- match_df(no_species, historic_summary)

```



sjekk hvilke arter som ligger utenom mine koordinater. Er det noen arter som er mer sannsynelig å finne midt i fjorden?


fins det noken sekvenser som er fisk men ikkje får assignment - asv, lowest common ancestor


```{r}


df_historic <- cc
df_lone <- original

df_lone$type <- "eDNA method"
df_historic$type <- "Traditional method"

df_historic <- unique(df_historic)

#hist(as.numeric(df_historic$out), breaks = 80)

df_both_map <- rbind(df_lone,df_historic)

#	Let's use Taurulus_bubalis as an example species. You can repeat this same code for all the species that you are interested in by just assigning a different species below:

genus_species <- "Gadus_morhua"

# Here you subset the data for only the species of interest
df_genus_species <- subset(df_both_map,species==genus_species)

#Here you create a dataframe that counts the number of observations at each sampling point
df_genus_species2 <- df_genus_species %>%
  group_by(out) %>%
  dplyr::summarize(n= length(out))

# Here you merge the "species of intersest" data frame and the number of observations at each sampling point
df_genus_species <- merge(df_genus_species,df_genus_species2,by="out")

# Here you create a new categorical variable indicating whether the species of interest was present in both the historical record and your sampling, or only historical record or your sampling.
df_genus_species$presence <- ifelse(df_genus_species$n>1,"Present in both",df_genus_species$type)

# Here you create an index dataframe for all those sampling sites where this species was present
sites_species <- as.data.frame(unique(df_genus_species$out))
colnames(sites_species)[1] ="out"

#Here you create an index dataframe for all those sampling sites where this species was absent
all_sites <- as.data.frame(1:79)
colnames(all_sites)[1] ="out"
sites_empty <- rbind(sites_species,all_sites)



#sites_empty <- sites_empty %>%
  #group_by(out) %>%
  #mutate(out2 = ifelse(length(out)>1,1,0))

sites_empty <- as.data.frame(table(sites_empty))



sites_empty <- subset(sites_empty,Freq==1)
sites_empty$species <- genus_species
sites_empty$type <- NA
sites_empty$n <- 0
sites_empty$presence <-"Absent" # Here you finally assign "absent" value to those sites where this species was absent in both the historical record and your data.
sites_empty <- sites_empty[,-2]

#Here you make the final dataframe that you can use in your maps.

df_genus_species <- rbind(df_genus_species,sites_empty)

df_genus_species <- df_genus_species[,c(1,2,5)]

df_genus_species <- df_genus_species[!duplicated(df_genus_species), ]

#rownames(df_genus_species) <- df_genus_species$out)

coordinates <- subset(data_cleaned, select = c(Lat,Lon))
coordinates$out <- 1:79


df_genus_species <- merge(df_genus_species, coordinates, by = "out")

set_shapes <- setNames(c(18, 15, 17, 16), levels(unique(df_genus_species$presence)))
# Now you can just add the coordinates for each sampling site and make your map figure. In order to get the different symbols to represent the different presence status, just add "presence" column as a grouping variable.
png("cod-map.png", height = 500, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df_genus_species, aes(x=Lon, y=Lat, color = presence, shape=presence)) +
  scale_shape_manual(values = set_shapes) +
  labs(x = "Longitude", y = "Latitude")+
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()
```
```{r}
genus_species <- "Ctenolabrus_rupestris"

# Here you subset the data for only the species of interest
df_genus_species <- subset(df_both_map,species==genus_species)

#Here you create a dataframe that counts the number of observations at each sampling point
df_genus_species2 <- df_genus_species %>%
  group_by(out) %>%
  dplyr::summarize(n= length(out))

# Here you merge the "species of intersest" data frame and the number of observations at each sampling point
df_genus_species <- merge(df_genus_species,df_genus_species2,by="out")

# Here you create a new categorical variable indicating whether the species of interest was present in both the historical record and your sampling, or only historical record or your sampling.
df_genus_species$presence <- ifelse(df_genus_species$n>1,"Present in both",df_genus_species$type)

# Here you create an index dataframe for all those sampling sites where this species was present
sites_species <- as.data.frame(unique(df_genus_species$out))
colnames(sites_species)[1] ="out"

#Here you create an index dataframe for all those sampling sites where this species was absent
all_sites <- as.data.frame(1:79)
colnames(all_sites)[1] ="out"
sites_empty <- rbind(sites_species,all_sites)


#sites_empty <- sites_empty %>%
  #group_by(out) %>%
  #mutate(out2 = ifelse(length(out)>1,1,0))

sites_empty <- as.data.frame(table(sites_empty))



sites_empty <- subset(sites_empty,Freq==1)
sites_empty$species <- genus_species
sites_empty$type <- NA
sites_empty$n <- 0
sites_empty$presence <-"Absent" # Here you finally assign "absent" value to those sites where this species was absent in both the historical record and your data.
sites_empty <- sites_empty[,-2]

#Here you make the final dataframe that you can use in your maps.

df_genus_species <- rbind(df_genus_species,sites_empty)

df_genus_species <- df_genus_species[,c(1,2,5)]

df_genus_species <- df_genus_species[!duplicated(df_genus_species), ]

#rownames(df_genus_species) <- as.numeric(df_genus_species$out)

coordinates <- subset(data_cleaned, select = c(Lat,Lon))
coordinates$out <- 1:79


df_genus_species <- merge(df_genus_species, coordinates, by = "out")

# Now you can just add the coordinates for each sampling site and make your map figure. In order to get the different symbols to represent the different presence status, just add "presence" column as a grouping variable.
png("bergnebb-map.png", height = 500, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df_genus_species, aes(x=Lon, y=Lat, color = presence, shape=presence)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = set_shapes) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()
```
```{r}


df2 <- subset(df_both_map, species %in% c("Oncorhynchus_gorbuscha", "Cyprinus_carpio", "Tinca_tinca", "Carassius_auratus"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
png("invasive-map.png", height = 400, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = c(15, 16)) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()

df2 <- subset(df_both_map, species %in% c("Engraulis_encrasicolus", "Scomber_scombrus", "Entelurus_aequoreus", "Trisopterus_minutus", "Sardina_pilchardus"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
png("warm-water-map.png", height = 400, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = c(15, 16)) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()
```
```{r}
# Basic piechart, må fikse
df <- as.data.frame(table(historical.2000$institution))

colnames(df) <- c("Institution", "frequency")

#eller simpelthen bare
png("org_barplot.png", height = 300, width = 450)
ggplot(df, aes(x="", y=frequency, fill=Institution)) +
  theme_classic() + 
 geom_bar(stat="identity")
dev.off()
```
```{r}
df2 <- subset(df_both_map, species %in% c("Buglossidium_luteum", "Ciliata_septentrionalis"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
png("novel-map.png", height = 400, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=species)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = c(15, 16)) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()
```

```{r}

df2rare <- subset(df_both_map, species %in% c("Trisopterus_luscus", "Phrynorhombus_norvegicus", "Ammodytes_marinus", "Raniceps_raninus", "Centrolabrus_exoletus", "Enchelyopus_cimbrius"))
df2rare$species <- factor(df2rare$species)

df3 <- merge(df2rare, coordinates)
png("rare-map.png", height = 500, width = 450)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = c(15, 16)) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()

df2 <- subset(df_both_map, species %in% refound$species)
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
png("refound-map.png", height = 400, width = 400)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species)) +
  labs(x = "Longitude", y = "Latitude")+
  scale_shape_manual(values = c(15, 16)) +
  theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))
dev.off()

 df2 <- subset(df_both_map, species %in% c("Anguilla_anguilla"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
#ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
# geom_point(data=df3, aes(x=Lon, y=Lat, color = type, shape=type)) +
#  labs(x = "Longitude", y = "Latitude")
```
```{r}



 df2 <- subset(df_both_map, species %in% c("Gymnocephalus_cernua", "Entelurus_aequoreus", "Abramis_brama", "Phrynorhombus_norvegicus"))
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")

 df2 <- subset(df_both_map, species %in% bbee$species)
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")
```




```{r}

historic <- subset(historical30, select = c(Lon, Lat))
colnames(historic) <- c("Lon", "Lat")
historic_base <- rbind(base.df, historic)

species30 <- unique(historical30$species)

historic.mtx <- as.matrix(historic)


hx <- lapply(1:nrow(historic), function(x){
    spDistsN1(pts = base.mtx, pt = historic.mtx[x, ], longlat =T)}) %>%
lapply(which.min) %>%
unlist %>%
bind_cols(historic, out = .)

hy <- lapply(1:nrow(historic), function(x){
    spDistsN1(pts = base.mtx, pt = historic.mtx[x, ], longlat =T)}) %>%
lapply(min) %>%
unlist %>%
bind_cols(historic, distance = .)

hw <- cbind(hx, hy)
hw <- hw[, !duplicated(colnames(hw))]
hw <- cbind(hw, historical30$species)
hw <- filter(hw, distance<1)

colnames(hw) <- c("Lon", "Lat", "out", "distance", "species")
hc <- subset(hw, select = c(out, species))
out <- hc$out

hc <- gsub(" ", "_", hc$species)
hc <- as.data.frame(cbind(hc,out))
colnames(hc) <- c("species", "out")


df_lone$type <- "lones_data"
hc$type <- "historic"

df_historic <- unique(hc)

#hist(as.numeric(df_historic$out), breaks = 80)

df_both_map_old <- rbind(df_lone,hc)


df2 <- subset(df_both_map_old, species %in% refound$species)
df2$species <- factor(df2$species)

df3 <- merge(df2, coordinates)
ggmap(get_stamenmap(c(10, 58.8, 11, 60),  maptype = "toner-lite", zoom = 10)) +
  geom_point(data=df3, aes(x=Lon, y=Lat, color = species, shape=type)) +
  labs(x = "Longitude", y = "Latitude")
```
kanskje eg må passe på at dei ikkje overlapper




```{r}

#Ditta går ikkje når 
c_species <- gsub(" ", "_", cc$species)
c_species <- unique(c_species)
o_species <- as.character(unique(original$species))
species_name <- c(c_species, o_species)
species_name <- unique(species_name)


his2017 <- subset(historical.2000, select = c("validScientificName", "longitude", "latitude"))
his2017$validScientificName <- gsub(" ", "_", his2017$validScientificName)
colnames(his2017) <- c("Species", "Lon", "Lat")





library(reshape2)

species_name <- colnames(species_cleaned)
test3 = match_df(cc, original, on = "out")
test4 = match_df(original, cc, on = "out")

mx3 <- t(as.matrix(with(test3, table(species, out)))) > 0
mx4 <- t(as.matrix(with(test4, table(species, out)))) > 0

mx5 <- mx4[, colSums(mx4) > 0]

combined = cbind(mx3, mx5)
sn <- colnames(combined)

for (i in 1:length(sn)){
for (y in 1:length(sn)){
  if((y-i) >0){g <- (y-i)}
  else{}
  if (i==y){
  }
  else{
  ct <- chisq.test(combined[,i], combined[,g], simulate.p.value = T)
  ft <- fisher.test(combined[,i], combined[,g], simulate.p.value = T)
  if (ct$p.value <0.05) {
    print(ct)
    print(i)
    print(y)
  }
  else{}
  if (ft$p.value <0.05) {
    print(ft)
    print(i)
    print(y)
  }
  else{}
  }
}
}
  


library(Matrix)
image(Matrix(mx))


```







```{r}
i.p <- c()
g.p <- c()
ft.p <- c()
for (i in 1:length(sn)){
for (y in 1:length(sn)){
  if((y-i) >0){g <- (y-i)}
  else{}
  if (i==y){
  }
  else{
  ft <- fisher.test(combined[,i], combined[,g], simulate.p.value = T)
  ft.p <- append(ft.p, ft$p.value)
  ft.p <- as.data.frame(ft.p)
  i.p <- append(i.p,i)

  g.p <- append(g.p, g)

  }
}
}


fdr <- p.adjust(ft.p, method = "fdr")
pval <- as.data.frame( ft.p)
sort(fdr)

for (f in 1:length(fdr)){
  if (fdr[f] < 0.05){
    fd <- append(fd, fdr[f])
    print(fdr[f])
    print(f)
  }
  else{}
}

for (y in 1:length(fd)) {
for (i in 1:length(ft.p)){
  if (ft.p[i] == fd[y]){
    print(i)
  }
  else{}
}
}


sort(p.adjust(ct.p, method = "fdr"))

x


```

```{r}
mx3 = cc
mx4 = original
sn <- unique(cc$species)
ft.p <- c()
ct.p <- c()

for (specie in sn){
  subset_cc = mx3[mx3$species == specie,]
  subset_og = mx4[mx4$species == specie,]
  merged_species = merge(subset_cc, subset_og, by = "out", all = TRUE)
  merged_species$species.y = as.character(merged_species$species.y)
  merged_species[is.na(merged_species)] <- 0
  if(length(unique(merged_species$species.x))>1 &&
     length(unique(merged_species$species.y))>1){

    ct <- chisq.test(merged_species$species.x, merged_species$species.y, simulate.p.value = T)
    ft <- fisher.test(merged_species$species.x, merged_species$species.y, simulate.p.value = T)
    ft.p <- append(ft.p, ft$p.value)
    ct.p <- append(ct.p, ct$p.value)
    }
  else{
    ft.p <- append(ft.p, NA)
    ct.p <- append(ct.p, NA)
  }  
}
names(ft.p) = sn
names(ct.p) = sn



ft.p.adjust <- p.adjust(ft.p, method = "hochberg", n = length(ft.p))
ct.p.adjust <- p.adjust(ct.p, method = "hochberg", n = length(ct.p))

ft.p.adjust <- as.data.frame(ft.p.adjust)
ct.p.adjust <- as.data.frame(ct.p.adjust)

both.test <- as.data.frame(cbind(ft.p.adjust, ct.p.adjust))
both.test$species  <- rownames(both.test)
colnames(both.test) <- c("Fisher test","Chi-squared test", "species")

both.test <- na.omit(both.test)

both.test <- melt(both.test)

# Create a binary variable indicating whether each value is below 0.1
both.test$below_005 <- ifelse(both.test$value < 0.05, "below 0.05", ifelse(both.test$value < 0.1, "below 0.1", "above 0.1"))

# Plot the heatmap with a manual fill scale

my_breaks = c(0.01, 0.05, 0.001, 0.0001)
ggplot(both.test, aes(variable, species, fill = below_005)) +
  geom_tile() +
  scale_fill_manual(values = c("dark blue", "blue", "light blue"), 
                    labels = c("above 0.1", "below 0.1", "below 0.05"))


png("fishertest.png", height = 700, width = 700)
ggplot(both.test, aes(variable, species, fill = value)) +  geom_tile() + scale_fill_gradientn(colors = c("blue", "white", "yellow","orange", "red"), breaks = my_breaks, trans = "log", labels = c("0.01", "0.05", "0.001", "0.0001"))+
  theme_classic() + 
    theme(axis.text = element_text(size = 14),
        axis.title.x = element_text(size = 16), # set size of x axis label
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16))

dev.off()



```



```{r}

library(ggplot2)
library(VennDiagram)

# Create a new variable that indicates whether the species is below or above a certain latitude
lat_group <- ifelse(data_cleaned$Lat > 59.667800, "above", "below")

df4 <- data.frame(lat_group, species_cleaned)

melted_df= melt(df4)
melt_df <- filter(melted_df, value !=0)
melt_df <- subset(melt_df, select = -c(value))

above = c()
below = c()
for (i in 1:length(melt_df$lat_group)){
  if (melt_df$lat_group[i] =="above"){
  above <- melt_df$variable[i]
  }
  else{below <- melt_df$variable[i]}
}




above <- as.data.frame(A$`Inner Oslo fjord`)
above <- unique(above)
colnames(above) <- "species"
below <- as.data.frame(A$`Outer Oslo fjord`)
below <- unique(below)
colnames(below) <- "species"

bee <- rbind(above, below)
bee <- as.data.frame(table(bee))
bee <- subset(bee,Freq==1)
bee <- as.data.frame(bee)
colnames(bee) <- c("species", "v1")

library(plyr)

abee <- match_df(above, bee, on = "species")
bbee <- match_df(below, bee, on = "species")

spge <- rbind(abee, bbee)

library("ggvenn")
  
# use list as input
A <-list("Outer Oslo fjord" = melt_df[melt_df$lat_group == "below", "variable"],
           "Inner Oslo fjord" = melt_df[melt_df$lat_group == "above", "variable"])
  
# create venn diagram and display all the sets
png("venndiagram.png", height = 500, width = 650)
ggvenn(A) +
     theme(
        legend.text = element_text(size = 14), # set size of legend text
        legend.title = element_text(size = 16),
        plot.subtitle = element_text(size = 12), 
        text = element_text(size = 46)) # set the font size to 18
dev.off()
```


```{r}
mysp <- colnames(species_cleaned)

A <-list("species in the 30s" = species30,
          "my species"= mysp)
  
# create venn diagram and display all the sets
ggvenn(A)


historical70 <- historic.data[time > as.Date("1970-1-1"), ]
historical70 <- historical70[time < as.Date("1990-1-1"), ]
species70 <- unique(historical70$species)

A <-list("species in the 70s" = species70,
          "my species"= mysp)
  
# create venn diagram and display all the sets
ggvenn(A)

```

